### 2022-02-07 manuscript/manuscript.Rmd
Add references

Bone is known to increase its tissue mass and enhance its micro-architecture arrangement in response to mechanical stimulation, particularly at higher magnitudes. This type of high-intensity impact activies have shown to be more associated with bone health improval, even if they are usually performed in low quantities throughout the day. In fact, Vainionpää et al demonstrated that performing a relatively low number (<100) of pACC above 4.9*g* daily is positively associated with increases in bone mineral density. Usually, this acceleration magnitude can be achieved mostly by ballistic movements such as jumps. Nevertheless, accelerometry-based mechanical loading regression models have mostly been validated with a set of activies, such as walking and running, that elict relatively low acceleration magnitudes, usually ranging from approximately 1.2 to 6*g* [@Neugebauer2014; @Veras2020].

The use of a prediction model based on accelerometer data should ideally be done in conditions similar to the study that validated the model, specially in terms of sample characteristics and type and intensity of the activities performed. Therefore, if we want to accurately predict the mechanical loading caused by high-impact activities, we should choose a prediction model that has been validated using these activities. Even though the cross-validation accuracy of such a model was poorer compared with other models developed for activities such as walking and running, as demonstrated by our results, the use of these models is crucial to achieve an acceptable prediction accuracy. It was demonstrated by testing a model previously developed by our group [@Veras2020] in a similar sample, but on walking, on the accelerometer-recorded jumping data from the current study. The accuracy indices yielded by applying these models have greatly worsened and these equations also produced some abnormal values, mainly in the pGRF model, in which it predicted negative values of force. These negative pGRF values were mainly as a result of a predictor, the second degree polynomial of the acceleration, possessing a negative coefficient. While this coefficient captured the plateauing of pGRF as pACC magnitude augments during walking, effectively improving the model [@Veras2020] for walking, this same pattern is not present in the pGRF vs. pACC relationship during jumping, and, therefore, only impairs the accuracy of these equations when applyed to this kind of activity.

Loading rate is a highly sensitive parameter, that is usually calculated by deriving the force from the beggining of the foot contact to the floor to the point where the pGRF is achieved. This is a small time window, usually less than 0.2 seconds for the jumps that were performed in our protocol, and requires a fine sampling acquisition that captures small graduations on the force values throughout time, which is difficult to achieve with common accelerometer sampling frequecies, that are usually 100Hz at maximum. Based on this and on on results from previous investigations [@Fortune2014; @Veras2020], pLR prediction models are expected to have a lower accuracy compared to the pGRF equations. Nevertheless, although our pLR prediction models had a worse accuracy compared with the pGRF models, its accuracy was similar or even better than the results from previous studies that developed models to predict pLR during walking only [@Fortune2014; @Veras2020].

In our protocol, three accelerometer placements (ankle, lower back and hip) and two vectors (resultant and vertical) have been tested. Overall, the prediction models performance was reasonably consistent among the different placement and vector combinations, however, some conclusions can be draw. First, the hip accelerometer placement, that has traditionally been widely employed in research for the direct measure of physical activity, yielded a sligthly better accuracy compared the other placements tested. This could be useful, as previous data collected from hip-worn accelerometers, even for other purposes, can be adequately used to obtain mechanical loading predictions. Also, studies that aim to extract mechanical loading and other variables from accelerometry data, that use the hip as the default placement, can comfortably do so. Second, as the magnitude of the resultant vector of both pGRF and pLR comes from (?) the vertical vector, and there were no important (?) differences in performance between models using both vectors, they can be used interchangeably to predict mechanical loading. Besides that, while models using data from the vertical vector allow the use uniaxial accelerometers, data derived from the resultant vector have the important advantage of not depending on the correct acceleromter axes orientation, that cannot be guaranteed in free-living data collection.

The development of accelerometry-based mechanical loading prediction models was done with the direct purpose of allowing to assess the loading beared by the skeletal system during daily physical activity. This would allow some clinical applications. For instance, these models, when used in combination with other prediction equations developed for lower impact activities [@Veras2020], can be used to fully analyse the mechanical loading spectrum that people with different characteristics are submitted to daily and relate to bone health parameters. This would allow a more complete understanding on how mechanical loading influence bone healh, and also allow to identify which loading volume and magnitude range, for example, are more efficient in inducing bone health benefits. Also, by the development and distribution of an R package that automates the steps necessary to apply the models that were proposed was made to enhance its applicability. This can prove to be an important tool, as the application of these models depend on non-trivial data processing skills, such as applying digital filters to reduce noise from the signal, detection of the peaks in this signal and the computation of derivatives to obtain the acceleration rate, necessary for the prediction of pLR.

>>>>

Bone is known to increase its tissue mass and enhance its micro-architecture arrangement in response to mechanical stimulation [@Turner2003; @Fonseca2014], particularly at higher magnitudes [@Vainionpaa2006; @Stiles2017]. This type of high-intensity impact activies have shown to be more associated with bone health improval, even if they are usually performed in low quantities throughout the day. In fact, Vainionpää et al [@Vainionpaa2006] demonstrated that performing a relatively low number (<100) of pACC above 4.9*g* daily is positively associated with increases in bone mineral density. Usually, this acceleration magnitude can be achieved mostly by ballistic movements such as jumps [@Vainionpaa2006; @Rowlands2012]. Nevertheless, accelerometry-based mechanical loading regression models have mostly been validated with a set of activies, such as walking and running, that elict relatively low acceleration magnitudes, usually ranging from approximately 1.2 to 6*g* [@Neugebauer2014; @Veras2020].

The use of a prediction model based on accelerometer data should ideally be done in conditions similar to the study that validated the model, specially in terms of sample characteristics and type and intensity of the activities performed [@Welk2005]. Therefore, if we want to accurately predict the mechanical loading caused by high-impact activities, we should choose a prediction model that has been validated using these activities. Even though the cross-validation accuracy of such a model was poorer compared with other models developed for activities such as walking and running, as demonstrated by our results, the use of these models is crucial to achieve an acceptable prediction accuracy. It was demonstrated by testing a model previously developed by our group [@Veras2020] in a similar sample, but on walking, on the accelerometer-recorded jumping data from the current study. The accuracy indices yielded by applying these models have greatly worsened and these equations also produced some abnormal values, mainly in the pGRF model, in which it predicted negative values of force. These negative pGRF values were mainly as a result of a predictor, the second degree polynomial of the acceleration, possessing a negative coefficient. While this coefficient captured the plateauing of pGRF as pACC magnitude augments during walking, effectively improving the model [@Veras2020] for walking, this same pattern is not present in the pGRF vs. pACC relationship during jumping, and, therefore, only impairs the accuracy of these equations when applyed to this kind of activity.

Loading rate is a highly sensitive parameter, that is usually calculated by deriving the force from the beggining of the foot contact to the floor to the point where the pGRF is achieved [@Liikavainio2007; @Fortune2014; @Veras2020]. This is a small time window, usually less than 0.2 seconds for the jumps that were performed in our protocol, and requires a fine sampling acquisition that captures small graduations on the force values throughout time, which is difficult to achieve with common accelerometer sampling frequecies, that are usually 100Hz at maximum. Based on this and on on results from previous investigations [@Fortune2014; @Veras2020], pLR prediction models are expected to have a lower accuracy compared to the pGRF equations. Nevertheless, although our pLR prediction models had a worse accuracy compared with the pGRF models, its accuracy was similar or even better than the results from previous studies that developed models to predict pLR during walking only [@Fortune2014; @Veras2020].

In our protocol, three accelerometer placements (ankle, lower back and hip) and two vectors (resultant and vertical) have been tested. Overall, the prediction models performance was reasonably consistent among the different placement and vector combinations, however, some conclusions can be draw. First, the hip accelerometer placement, that has traditionally been widely employed in research for the direct measure of physical activity [@Migueles2017], yielded a sligthly better accuracy compared the other placements tested. This could be useful, as previous data collected from hip-worn accelerometers, even for other purposes, can be adequately used to obtain mechanical loading predictions. Also, studies that aim to extract mechanical loading and other variables from accelerometry data, that use the hip as the default placement, can comfortably do so. Second, as the magnitude of the resultant vector of both pGRF and pLR, in the activities that were tested, is derived mostly from the vertical vector, and there were no considerable differences in performance between models using both vectors, they can be used interchangeably to predict mechanical loading. Besides that, while models using data from the vertical vector allow the use uniaxial accelerometers, data derived from the resultant vector have the important advantage of not depending on the correct acceleromter axes orientation, that cannot be guaranteed in free-living data collection [@Welk2005].

The development of accelerometry-based mechanical loading prediction models was done with the direct purpose of allowing to assess the loading beared by the skeletal system during daily physical activity. This would allow some clinical applications. For instance, these models, when used in combination with other prediction equations developed for lower impact activities [@Veras2020], can be used to fully analyse the mechanical loading spectrum that people with different characteristics are submitted to daily and relate to bone health parameters. This would allow a more complete understanding on how mechanical loading influence bone healh, and also allow to identify which loading volume and magnitude range, for example, are more efficient in inducing bone health benefits. Also, by the development and distribution of an R package that automates the steps necessary to apply the models that were proposed was made to enhance its applicability [@Veras2021]. This can prove to be an important tool, as the application of these models depend on non-trivial data processing skills, such as applying digital filters to reduce noise from the signal, detection of the peaks in this signal and the computation of derivatives to obtain the acceleration rate, necessary for the prediction of pLR.
